\mytitle

### 2015-07-31 MSc_diss.tex
Correcting the structure

According to the Intergovernmental Panel on Climate Change (IPCC) Fifth Assessment report (AR5) for present conditions (around year 2000) the global mean tropospheric ozone budget is approximately 331 Tg \citep{Myhre2013}. However, despite general agreement on how the drivers impact ozone concentrations at global and regional scales, its budget varies considerably between different modelling and observational studies \citep{Stevenson2006,Wild2007,Young2012}. Table \ref{tab:O3budget} demonstrates that the range of the global tropospheric ozone budget estimates is approximately 60 Tg which is 18\% of the mean. In terms of ozone production the variation is even higher and account for 28\% of the corresponding value. In that regard, two questions arise: why does the tropospheric ozone budget differ between models and observations? And secondly, why does it vary to this extent, especially in terms of production? The only way to answer these questions is to perform a rigorous investigation of the factors that drive tropospheric ozone and attribute those differences to respective factors \citep{Young2012}.

@@

The second way to understanding the discrepancies between models and observations is to conduct an idealised theoretical study using a conceptual model. The advantage of this approach is that such kind of models can be quite easily constructed, and more importantly can be focused on representation of a certain chemical reactions which presumably could explain the discrepancy between models and observations. An example of such set of reactions is alkyl nitrate chemistry.

\subsection{Alkyl nitrates}
\subsubsection{Alkyl nitrates chemistry}
Alkyl nitrates are important tropospheric trace gases which are formed following the same chemistry that leads to the production of ozone \citep{Reeves2007}. The simplest example, in case of $CH_4$ oxidation, can be described by the set of reactions below (also shown in Figure 2).

%\input{O3CycleScheme}

In the presence of UV light shorter than 320 nm and water vapour $O_3$ photolysis is the main source of OH radicals in the troposphere:
\begin{equation} \label{eq:O3hv}
O_3 + hv \rightarrow O(^1D) + O_2
\end{equation}
\begin{equation} \label{eq:O1D+H2O}
O(^1D) + H_2O \rightarrow 2OH
\end{equation}
Once formed the OH radicals react mainly with $CO$ and $CH_4$ to initiate the $O_3$ formation or removal cycles by producing peroxy radicals ($HO_2$, $CH_3O_2$):
\begin{equation} \label{eq:OH+CO}
OH + CO \rightarrow HO_2 + CO_2
\end{equation}
\begin{equation} \label{eq:OH+CH4}
OH + CH_4 \rightarrow CH_3O_2 + H_2O
\end{equation}

The fate of the peroxy radicals depends on the presence of NOx. At low NOx levels (remote regions), the $HO_2$ radical reacts with another HO2 to form hydrogen peroxide, H2O2, or with the methyl peroxy radical, $CH_3O_2$, to form methyl hydrogen peroxide ($CH_3OOH$):
\begin{equation} \label{eq:HO2+HO2}
HO_2 + HO_2 \rightarrow H_2O_2 + O_2
\end{equation}
\begin{equation} \label{eq:HO2+CH3O2}
HO_2 + CH_3O_2 \rightarrow CH_3OOH + O_2
\end{equation}
On a whole, at low NOx levels a net $O_3$ loss takes place, because the reaction sequence is initiated by $O_3$ photolysis. Plus under the same conditions, some additional $O_3$ removal occurs due to the re-action of $HO_2$ radicals with $O_3$,
\begin{equation} \label{eq:HO2+O3}
HO_2 + O_3 \rightarrow OH + 2O_2
\end{equation}
leading to the regeneration of OH radicals as part of an $O_3$ depleting HOx catalytic cycle.

@@

>>>>

According to the Intergovernmental Panel on Climate Change (IPCC) Fifth Assessment report (AR5) for present conditions (around year 2000) the global mean tropospheric ozone budget is approximately 331 Tg \citep{Myhre2013}. However, despite general agreement on how the drivers impact ozone concentrations at global and regional scales, its budget varies considerably between different modelling and observational studies \citep{Stevenson2006,Wild2007,Young2012}. Table \ref{tab:O3budget} demonstrates that the range of the global tropospheric ozone budget estimates is approximately 60 Tg which is 18\% of the mean. In terms of ozone production the variation is even higher and accounts for 28\% of the corresponding value. In that regard, two questions arise: why does the tropospheric ozone budget differ between models and observations? And secondly, why does it vary to this extent, especially in terms of production? The only way to answer these questions is to perform a rigorous investigation of the factors that drive tropospheric ozone and attribute those differences to respective factors \citep{Young2012}.

@@

! add something from report!

The second way to understanding the discrepancies between models and observations is to conduct an idealised theoretical study using a conceptual model. The advantage of this approach is that such kind of models can be quite easily constructed, and more importantly can be focused on representation of a certain chemical reactions which presumably could explain the discrepancy between models and observations.

\subsubsection*{Ozone tropospheric chemistry}
%\input{O3CycleScheme}
Having clarified the role of tropospheric ozone, uncertainties in its budget and having identified the ways of revealing the cause of these uncertainties, it is time to remind ourselves the chemistry that controls the abundance of ozone in the troposphere. The key chemical processes that lead to $O_3$ production and destruction are driven by reaction cycles involving free-radical intermediates, which are mainly formed from the photolysis of $O_3$ itself (\ref{reac:O3+hv}). In the presence of UV light shorter than 320 nm and water vapour $O_3$ photolysis is the main source of OH radicals in the troposphere:
\begin{equation} \label{reac:O3+hv}
O_3 + hv \rightarrow O(^1D) + O_2
\end{equation}
\begin{equation} \label{eq:O1D+H2O}
O(^1D) + H_2O \rightarrow 2OH
\end{equation}
Once formed the OH radicals react mainly with $CO$ and $CH_4$ to initiate the $O_3$ formation or removal cycles by producing the peroxy radicals ($HO_2$, $CH_3O_2$):
\begin{equation} \label{eq:OH+CO}
OH + CO \rightarrow HO_2 + CO_2
\end{equation}
\begin{equation} \label{eq:OH+CH4}
OH + CH_4 \rightarrow CH_3O_2 + H_2O
\end{equation}

The fate of the peroxy radicals depends on the presence of $NO_x$ ($=NO + NO_2$). At low $NO_x$ levels (remote regions), the $HO_2$ radical reacts with another $HO_2$ to form hydrogen peroxide ($H_2O_2$) or reacts with the methyl peroxy radical ($CH_3O_2$) to form methyl hydrogen peroxide ($CH_3OOH$):
\begin{equation} \label{eq:HO2+HO2}
HO_2 + HO_2 \rightarrow H_2O_2 + O_2
\end{equation}
\begin{equation} \label{eq:HO2+CH3O2}
HO_2 + CH_3O_2 \rightarrow CH_3OOH + O_2
\end{equation}
On a whole, at low $NO_x$ levels a net $O_3$ loss takes place, because the reaction sequence is initiated by $O_3$ photolysis. Plus under the same conditions, some additional $O_3$ removal occurs due to the reaction of $HO_2$ radicals with $O_3$,
\begin{equation} \label{eq:HO2+O3}
HO_2 + O_3 \rightarrow OH + 2O_2
\end{equation}
leading to the regeneration of OH radicals as part of an $O_3$ depleting $OH-HO_2$ ($HO_x$) catalytic cycle.

@@

\subsection{Alkyl nitrates}
\subsubsection{Alkyl nitrates chemistry}
(An example of such set of reactions is alkyl nitrate chemistry)

Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases, which   due to one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}.
